# R Package- anatomy of an R package

```{r Load libraries}
library(devtools)
library(fs)
```

## Folder structure

What are the files and folders in the package?

```{r Show folder structure}
dir_tree(all = TRUE, recurse = 1)
```

### `.gitignore`

Git will not track some standard behind-scenes files created by R and RStudio.

### `.Rbuildignore`

`.Rbuildignore` It lists files and directories needed for development but not necessary to end users. It .

**These two "ignore" files have very different syntax!**

`.Rbuildignore` uses [Perl-like **regular expression**](https://cran.r-project.org/doc/manuals/R-exts.html), while `.gitignore` treats a pattern as [**shell glob** suitable for consumption by fnmatch(3) with the FNM_PATHNAME flag](https://www.git-scm.com/docs/gitignore/2.9.5).

Both files ignore the `*/.Rproj.user` folder.

-   `.gitignore`

".Rproj.user"

-   `.Rbuildignore`

"\^.Rproj.user\$"

With the `usethis` package, we can use `use_git_ignore` to write to `.gitignore`, and `use_build_ignore()` to write to `.Rbuildignore`.

```{r Add files that need to be ignored}
file.create("plan.Rmd") # For my personal use only

use_data_raw() # the data you include in data/ is a cleaned up version of raw data here, typically some data you have gathered from elsewhere. 
```

```{r `use_git_ignore` and `use_build_ignore`}
# Both .gitignore and .Rbuildignore files that you only want to keep locally
use_git_ignore("plan.Rmd")
use_build_ignore("plan.Rmd")

# .Rbuildignore Rmd file (but not the MD file)
use_build_ignore("README.Rmd")

# `use_data_raw()` already added "data_raw" to `.Rbuildignore`

```

### `DESCRIPTION`

Metadata about the package. Check with legal for company-Specific requirements!

-   Title: one-line description

-   Description: one-paragraph description

-   [Authors\@R](mailto:Authors@R){.email}: individuals or entities associated with the creation of the package, e.g. author, creator, copyright holder and funder.

-   License: mandatory and machine-readable.

```{r Use MIL license}
use_mit_license()
```


#### Add dependencies to `DESCRIPTION`

```{r Add import packages}
use_package("httr")
```

We will talk more about the difference between "Imports" and "Suggests". 

Don't do the following!

```{r Import tidyverse as dependency, error=TRUE}
use_package("tidyverse")
```

- Version: Remember to [bump version](https://github.com/tidyverse/dplyr/releases) before release

```
Package: apihealthcheck
Title: Check the health of API endpoints
Version: 0.0.0.9000
Authors@R: 
    c(person("Mark", "Wang", , "mark.wang@procogia.com", role = c("aut", "cre")), 
    person("Procogia", role = c("cph", "fnd")))
Description: Many R applications depend use API calls to retrieve and post data.
    This package helps check the health of API endpoints, thus enables timely
    detection and diagnosis of API errors. 
License: MIT + file LICENSE
Encoding: UTF-8
Roxygen: list(markdown = TRUE)
RoxygenNote: 7.2.3
```

### `NAMESPACE`

Functions exported for external use, or imported for internal use.

```{r Example of non-exported functions, error=TRUE}
library(remotes)

remotes:::is.named(mtcars) # Use ":::" to use non-exported function

remotes::is.named(mtcars)
```

### The `R/` directory

Functions are defined there.

### `regexcite.Rproj`

Make the package folder an R project. 