# R Package- anatomy of an R package

```{r Load libraries}
library(devtools)
library(fs)
```

## Folder structure

What are the files and folders in the package?

```{r Show folder structure}
dir_tree(all = TRUE, recurse = 1)
```

### `.gitignore`

Git will not track some standard behind-scenes files created by R and RStudio.

### `.Rbuildignore`

`.Rbuildignore` It lists files and directories needed for development but not necessary to end users. It .

**These two "ignore" files have very different syntax!**

`.Rbuildignore` uses [Perl-like **regular expression**](https://cran.r-project.org/doc/manuals/R-exts.html), while `.gitignore` treats a pattern as [**shell glob** suitable for consumption by fnmatch(3) with the FNM_PATHNAME flag](https://www.git-scm.com/docs/gitignore/2.9.5).

Both files ignore the `*/.Rproj.user` folder.

-   `.gitignore`

".Rproj.user"

-   `.Rbuildignore`

"\^.Rproj.user\$"

With the `usethis` package, we can use `use_git_ignore` to write to `.gitignore`, and `use_build_ignore()` to write to `.Rbuildignore`.

```{r Add files that need to be ignored}
file.create("plan.Rmd") # For my personal use only

use_data_raw() # the data you include in data/ is a cleaned up version of raw data here, typically some data you have gathered from elsewhere. 
```

```{r `use_git_ignore` and `use_build_ignore`}
# Both .gitignore and .Rbuildignore files that you only want to keep locally
use_git_ignore("plan.Rmd")
use_build_ignore("plan.Rmd")

# .Rbuildignore Rmd file (but not the MD file)
use_build_ignore("README.Rmd")

# `use_data_raw()` already added "data_raw" to `.Rbuildignore`

```

### `DESCRIPTION`

Metadata about the package. Check with legal for company-Specific requirements!

-   Title: one-line description

-   Description: one-paragraph description

-   [Authors\@R](mailto:Authors@R){.email}: individuals or entities associated with the creation of the package, e.g. author, creator, copyright holder and funder.

-   License: mandatory and machine-readable.

```{r Use MIL license}
use_mit_license()
```


#### Add dependencies to `DESCRIPTION`

```{r Add import packages}
use_package("httr")
```

We will talk more about the difference between "Imports" and "Suggests". 

Don't do the following!

```{r Import tidyverse as dependency, error=TRUE}
use_package("tidyverse")
```

- Version: Remember to [bump version](https://github.com/tidyverse/dplyr/releases) before release

```
Package: apihealthcheck
Title: Check the health of API endpoints
Version: 0.0.0.9000
Authors@R: 
    c(person("Mark", "Wang", , "mark.wang@procogia.com", role = c("aut", "cre")), 
    person("Procogia", role = c("cph", "fnd")))
Description: Many R applications depend use API calls to retrieve and post data.
    This package helps check the health of API endpoints, thus enables timely
    detection and diagnosis of API errors. 
License: MIT + file LICENSE
Encoding: UTF-8
Roxygen: list(markdown = TRUE)
RoxygenNote: 7.2.3
```

### `NAMESPACE`

Functions exported for external use, or imported for internal use.

```{r Example of non-exported functions, error=TRUE}
library(remotes)

remotes:::is.named(mtcars) # Use ":::" to use non-exported function

remotes::is.named(mtcars)
```

### The `R/` directory

Functions are defined there.

### `regexcite.Rproj`

Make the package folder an R project. 

## Write first functions

We will put the functions in `R/check_endpoint.R`:

```{r Create first function}
use_r("check_endpoint")
```

We define two functions there.

```
#' Get API endpoint status
#'
#' @param endpoint
#' An atomic character vector, the URL of API endpoint that needs to be checked
#'
#' @return
#' An atomic character vector, the status code of the API endpoint
#' @details
#' This function is not exported
#'
get_endpoint_status <- function(endpoint){

  config <- httr::authenticate(user = Sys.getenv("API_USER"), password = Sys.getenv("API_PASSWORD"), type = "basic")

  httr::GET(
    url = endpoint,
    config = config,
    handle = NULL,
    headers = httr::add_headers(
      .headers = c("Content-Type" = "application/json", "Accept" = "application/json")
    )
  ) |> httr::status_code()
}

#' Check the health of API endpoints
#'
#' @param endpoint_name
#' An atomic character vector, the name, saved as R environment variables
#' locally, of the API endpoint
#' @param response_time_limit
#' The maximum response time for the API endpoints, beyond which the endpoint
#' will be considered unhealthy.
#' @inheritParams get_endpoint_status
#' @export
#' @details
#' This function is exported for external use
#'
test_api_endpoint <- function(endpoint,
                              endpoint_name = NULL,
                              response_time_limit = 5) {
  if (!is.null(endpoint_name)) {
    endpoint <- Sys.getenv(endpoint_name)
  }

  endpoint_status <- R.utils::withTimeout(
    testthat::expect_no_error(get_endpoint_status(endpoint = endpoint)),
    timeout = response_time_limit ,
    onTimeout = "silent"
  )

  testthat::expect_true(!is.null(endpoint_status), label = glue::glue(
    "Endpoint {ifelse(is.null(endpoint_name), endpoint, endpoint_name)} took too long to respond."
  ))

  if (!is.null(endpoint_status)){
    testthat::expect_equal(
      endpoint_status,
      200,
      label = glue::glue(
        "Endpoint {ifelse(is.null(endpoint_name), endpoint, endpoint_name)} failed"
      )
    )
  }
}

```

```{r Import more packages}
use_package("R.utils")
use_package("glue")
use_package("testthat")
```

Load the function.

```{r Load function}
load_all()
test_api_endpoint("https://dummy.restapiexample.com/api/v1/employees")
```

Complete documentation.

```{r Complete documentation}
document()

?test_api_endpoint
```

The `test_api_endpoint` function should also have been added to the `NAMESPACE` file. 

## Make it a Shiny app

### Create Shiny app directory

```{r Make app firectory}
dir.create("inst")
dir.create("inst/app")
```
